---
title: "Variational Auto Encoder for Sequences"
output: html_notebook
---
```{r}
rm(list = ls())
```

```{r setup}
knitr::opts_knit$set(root.dir = '/home/harpo/hostdir/ia-dojo-repo/experiments/dga-gen/')
```

```{r message=FALSE, warning=FALSE}
library(keras)
library(tensorflow) 
```

# Load helpers functions
```{r message=FALSE, warning=FALSE}
source("code/R/functions/dgagen-helpers.R")
source("code/R/functions/dgagen-vae.R")
```
# Read domains
```{r}
library(urltools)
text <- readr::read_csv("rawdata/argencon.csv.gz")
text<-text %>% tidyr::separate(label,c("label","family"))
#text<-text %>% filter(grepl("normal",label))
#text<-text %>% group_by(family)%>% sample_frac(0.05)
text<-text %>% group_by(label)%>% sample_frac(0.1)
domains <-  text %>% pull(domain)
domains <- domains %>% urltools::host_extract()
domains <- domains  %>% tidyr::drop_na() %>% pull(host)
```
# Convert to one-hot encoding

```{r}
sentences_tokenized<-tokenize(domains %>% as.matrix, "n",maxlen)
shape <- c(nrow(sentences_tokenized$x), maxlen, length(valid_characters_vector))
seq_x<-to_onehot(sentences_tokenized$x,shape)
seq_x[1,,]
```

## Check one_hot
Check the one-hot implementation. Given a one-hot encoded domain, obtain the string

```{r}
dimen<-seq_x %>% dim()
enc<-""
for (i in 0:dimen[2]  ){
#      print(i)
      enc<-str_c(enc,valid_characters_vector[which.max(seq_x[1,i,])])
}
enc
```
 



# VAE
This is the current implementation of Variational auto-encoder for sequences.  The LSTM version is operational.

```{r}
latent_dim <-  64
vocab_size <-length(valid_characters_vector)
input_size <-  c(maxlen, vocab_size)
```


```{r}
vae_decoder<-create_vae_decoder(latent_dim = latent_dim, maxlen = maxlen,voc_size = vocab_size)
vae_encoder<-create_vae_encoder(latent_dim = latent_dim,input_size = input_size)
vae <- model_vae(vae_encoder, vae_decoder)
vae$summary()

```


# Training
```{r}
optimizer <- optimizer_adam(learning_rate = 0.005)
vae %>% compile(optimizer = optimizer)

```


```{r}
model_path <- "models/vael64best.keras/"
my_callbacks <- list(
  callback_model_checkpoint(model_path, 
                            save_best_only = TRUE)
)
vae %>% fit(seq_x,epochs = 5,
                batch_size = 512,
            validation_split=0.2)
```

## Save weights
```{r}
#vae$save("/tmp/keras-model")
#save_model_tf(vae,model_path)
#save_model_weights_tf(vae, "models/dgagen_vae__weights")
#load_model_weights_tf(vae, "models/dgagen_vae__weights")
```

#Testing

We pick a domain, encode it and decode it.

```{r}
# Pick domain
d<-domains[162100]
print(paste("domain: ",d))

# Preprocess
vectorized_test <- tokenize(d, "n", maxlen)
shape <- c(nrow(vectorized_test$x),
          maxlen,
          length(valid_characters_vector))
vectorized_test <- to_onehot(vectorized_test$x, shape)
dim(vectorized_test)
# Encode
encoded <- vae$encoder(vectorized_test)
encoded[[1]] %>% dim()
# Decode
decoded <- predict(vae$decoder, encoded[[1]], verbose = 0 )
dimen<-dim(decoded)
print(dimen)

# Print resulting domain
enc<-""
for (i in 1:dimen[2]  ){
  enc<-str_c(enc,valid_characters_vector[which.max(decoded[1,i,])])
}
enc<-substr(enc, start = 1,stop =nchar(d) )
print(paste("decoded domains:", enc))
```

## Generation

Given a domain generate small variations of it.

```{r message=FALSE}

d<-text %>% filter(label=="normal") %>% sample_n(10) %>% pull(domain) %>% urltools::host_extract() %>% tidyr::drop_na() %>% pull(host)
#did<-sample(1:100,10)
dga<-c()
for (j in 1:length(d)){
  #  d<-domains[j]
  vectorized_test <- tokenize(d[j], "n", maxlen)
  shape = c(nrow(vectorized_test$x),
            maxlen,
            length(valid_characters_vector))
  vectorized_test <- to_onehot(vectorized_test$x, shape)
  preds_encoder <- vae$encoder(vectorized_test)
  print(d[j])
  for (k in 1:10){
 
    eps<-rnorm(latent_dim,mean = 0,sd = 0.01)
    z <- tf$add(eps,preds_encoder[[1]])
    #z <- list(eps) %>% as_tensor()
    decoded <- predict(vae$decoder, z, verbose = 0 )
    dimen<-dim(decoded)
    enc<-""
    for (i in 1:dimen[2]  ){
      enc<-str_c(enc,valid_characters_vector[which.max(decoded[1,i,])])
    }
    enc<-substr(enc, start = 1,stop =nchar(d[j]) ) 
    dga<-c(dga,enc)
  }
}
dga %>% as.data.frame()

```


# DGA detector Results
Check the results of the generator against the CNN CACIC 2018 DGA detector.
```{r}
library(curl)
library(jsonlite)
res<-c()
reference<-factor(rep(1,length(dga)),levels=c(0,1))
for (j in seq(1:length(dga))){
  req<- curl::curl_fetch_memory(paste0("http://catanuso.duckdns.org:8000/predict?domain=",dga[j]))
  res[j]<-({jsonlite::parse_json(rawToChar(req$content))}$class)
}

data.frame(dga,res)
caret::confusionMatrix(data=as.factor(res),reference)
```
# Save the generated domains.
```{r}
dgadf<-data.frame(domains=dga)
readr::write_csv(dgadf,"/home/harpo/hostdir/ia-dojo-repo/experiments/dga-gen/data/generated-domains.csv")
```

```{bash eval=FALSE, include=FALSE}
harpo@joker:~/ia-dojo-repo/experiments/dga-gen/data$ docker run -i -v $PWD:/mnt registry.gitlab.com/cossas/dgad:4.1.1 client -f /mnt/generated-domains.csv -dc=domains >../results/dgad_detection.json

```

```{r}
library(jsonlite)
dgadres<-jsonlite::read_json("../../../results/dgad_detection.json")
dgadres<-dgadres %>% map( ~ifelse(.x$is_dga == TRUE, 1,0) ) %>% unlist()
caret::confusionMatrix(data=as.factor(dgadres),reference)
```



# Plots
## Encode domains into the embedded space.

```{r}

encoded_seq_x_vae <-predict(vae$encoder, seq_x)
encoded_seq_x_vae_mean<-encoded_seq_x_vae[[1]]
```

## Histogram
```{r}

library(ggplot2)
encoded_seq_x_vae_mean[,40] %>% as.data.frame() %>%
  ggplot()+
  geom_histogram(aes(x=.),fill='skyblue',color='white') +
  theme_classic()
```

## UMAP
```{r}
library(umap)
umap_data_vae <- umap(encoded_seq_x_vae_mean)
text_reduce_vae<-text %>% head(nrow(umap_data_vae$layout))
domains_data_umap_vae <- data.frame(umap_data_vae$layout, text_reduce_vae$family)
```


```{r fig.height=8, fig.width=8}
ggplot(domains_data_umap_vae %>% sample_frac(0.1)) +
  geom_point(aes(x=X1,y=X2,color=text_reduce_vae.family),alpha=0.9,size=2)+#,shape=1) +
  labs(subtitle = "Variational Autoencoder: Latent space 2D projection (UMAP) (zoomed-in view)",title="DOMAIN NAMES FAMILIES")+
  ylim(-5,5)+
  xlim(-6,5)+
  ylab("")+
  xlab("")+
  theme_classic()+
  ggdark::dark_theme_classic()
 # theme(legend.position = "none")

#plotly::ggplotly()

```
## PCA

```{r}
pca_data_vae<- prcomp(encoded_seq_x_vae_mean)
nrow(pca_data_vae$x)
text_reduce<-text %>% head(nrow(pca_data_vae$x))
domains_data_pca_vae <- data.frame(pca_data_vae$x, family=text_reduce$family)
```

```{r}
ggplot(domains_data_pca_vae %>% sample_frac(0.1)) +
  geom_point(aes(x=PC1,y=PC2,color=family),alpha=0.5,size=2) +
  theme(legend.position = NULL) +
  ylim(-5,5)+
  xlim(-5,5)+
  theme_classic()
```
